---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MIMIC4db

<!-- badges: start -->
<!-- badges: end -->

This work references data from the Medical Information Mart for Intensive Care MIMIC-IV database. 
MIMIC-IV is a large, freely-available database comprising de-identified health-related data from patients who
were admitted to the critical care units of the Beth Israel Deaconess Medical Center from 2008-2019.  Detailed
information can be obtained on the [MIMIC-IV website](https://mimic.mit.edu/docs/iv/).

MIMIC4db provides a tightly bound set of routines to reference the Google BigQuery version of the [MIMIC-IV v1.0](https://physionet.org/content/mimiciii/1.0/) database using 
base access routines from the [DBI](https://github.com/r-dbi/DBI) R package with an appropriate [bigrquery](https://github.com/r-dbi/bigrquery) DBIConnection. 

## Installation

You can install MIMIC4db from GitHub with:

``` r
# install.packages("devtools")
devtools::install_github("hdshea/MIMIC4db")
```

## Basic Usage

These are basic examples which show you how to solve a common problem.

Connecting to the database:

```{r pre_connect_internal, include=FALSE, eval=TRUE}
library(MIMIC4db)

con <- bigrquery::dbConnect(
    bigrquery::bigquery(),
    project = bigrquery::bq_test_project(),
    quiet = TRUE
)
```

```{r pre_connect, include=TRUE, eval=FALSE}
library(MIMIC4db)
# To connect to the database, you must provide an appropriate project (in the `"your-project"` slot below)
# which has billing set up and to which you have write access.  This project must be approved to access
# the PhysioNet BigQuery MIMIC schemas.

con <- bigrquery::dbConnect(
    bigrquery::bigquery(),
    project = "your-project",
    quiet = TRUE
)
```

Using the meta data function:

```{r meta_functions, include=TRUE}
core <- m4_meta_data()
core <- core[core$module == "mimic_core", ]
for(i in 1:length(core)) {
  cat(stringr::str_c("select * from physionet-data.", core[i,"module"], ".", core[i,"table"], "\n"))
}

neuro <- m4_service_decsriptions()
neuro <- neuro[startsWith(neuro$short_description, "Neurologic"), ]
neuro[,c("service", "short_description")]
```

Using the two base functions:

```{r base_functions, include=TRUE}
d_items <- m4_select_data(con, "select * from physionet-data.mimic_icu.d_items order by itemid limit 5")
d_items

d_items <- m4_get_from_table(con, "physionet-data.mimic_icu.d_items", where = "where itemid <= 220048")
d_items
```

Using the tables specific data access functions:

```{r table_functions, include=TRUE, eval=TRUE}
tab <- m4_chartevents_items(con)
tab[c(1:5), ]

pat <- m4_patients(con, cohort = 10137012)
pat

svs <- m4_services(con, cohort = 10137012)
svs[c(1:5), ]

evt <- m4_procedureevents(con, cohort = 12384098)
evt[c(1:5), ]
```

Disconnecting from the database:

```{r disconnect, include=TRUE, eval=TRUE}
bigrquery::dbDisconnect(con)
```


## Citations for MIMIC-IV publications:

Johnson, A., Bulgarelli, L., Pollard, T., Horng, S., Celi, L. A., & Mark, R. (2021). MIMIC-IV (version 1.0). PhysioNet. [https://doi.org/10.13026/s6n6-xd98](https://doi.org/10.13026/s6n6-xd98).

**PhysioNet**:
Goldberger, A., Amaral, L., Glass, L., Hausdorff, J., Ivanov, P. C., Mark, R., ... & Stanley, H. E. (2000).
PhysioBank, PhysioToolkit, and PhysioNet: Components of a new research resource for complex physiologic signals.
Circulation [Online]. 101 (23), pp. e215–e220.
